The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-09-04 JStatSoft_forestinventory_master-concordance.tex
update Andreas 04.09.2017(5)

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 105 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
1 97 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 211:%
121 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 221:%
1 88 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 310:%
132 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 322:%
1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 853:%
145}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 854:%
1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1373:%
147 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1385:%
1 42 1 1 2 11 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1450:%
160 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1460:%
1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1757:%
171 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1767:%
1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1998:%
182 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2007:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2012:%
192 11 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 105 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
1 97 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 211:%
121 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 221:%
1 87 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 309:%
132 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 321:%
1 126 1 1 2 4 0 1 2 14 1 1 5 7 0 1 2 2 1 1 2 24 0 1 2 4 1 1 2 4 0 1 2 2 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 157 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 5 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 852:%
145}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 853:%
1 119 1 1 2 8 0 1 2 4 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 8 0 1 2 193 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 11 0 1 3 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 7 1 1 7 10 0 1 3 2 1 1 2 12 0 1 2 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1372:%
147 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1384:%
1 42 1 1 2 11 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1449:%
160 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1459:%
1 20 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 18 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 4 1 1 2 8 0 1 2 22 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 39 1 1 3 5 0 1 2 23 1 1 3 5 0 1 2 19 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1756:%
171 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1766:%
1 20 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 16 1 1 4 1 1 1 2 11 0 1 2 18 1 1 5 7 0 1 2 4 1 1 3 5 0 1 2 10 1 1 2 4 0 1 2 9 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1997:%
182 8 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 2006:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 2011:%
192 11 1}

### 2017-09-04 JStatSoft_forestinventory_master.rnw
update Andreas for alex_version 2
update Andreas 04.09.2017(5)

%% an abstract and keywords
\Abstract{
Forest inventories provide reliable and evidence-based information to assess the state and development of forests over time.  They typically consist of a random sample of plot locations in the forest that are assessed individually by hand. Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs (e.g. LiDAR, stereophotogrammetry, etc.) is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision. With respect to this objective, the application of \textit{multiphase} forest inventory methods (e.g. double- and triple-sampling regression estimators) has proved to be efficient. While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent. The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling. The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata. This article summarizes the mathematical theory of this family of design-based estimators and demonstrates how they can be easily implemented in \proglang{R}.}

@@

% new commands
\newcommand{\bwi}{BWI3}
\newcommand{\var}{\mathbb{V}}

>>>>

%% an abstract and keywords
\Abstract{
Forest inventories provide reliable and evidence-based information to assess the state and development of forests over time.  They typically consist of a random sample of plot locations in the forest that are assessed individually by hand. Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs (e.g. LiDAR, stereophotogrammetry, etc.) is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision. With respect to this objective, the application of \textit{multiphase} forest inventory methods (e.g. double- and triple-sampling regression estimators) has proved to be efficient. While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent. The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling. The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata. This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in \proglang{R}.}

@@

% new commands
\newcommand{\bwi}{BWI3}
\newcommand{\var}{\mathbb{V}}
\newcommand{\EX}{\mathbb{E}}

### 2017-09-04 rnw/Introduction.Rnw
update Andreas 04.09.2017(5)

% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
Despite these promising methods, a standard application of two and three-phase sampling methods in forest practice has been hampered by a lack of available software. One exception is the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach. However, a more comprehensive software package covering a larger variety of sampling designs and estimators applicable to forest inventories has up to now been missing, which was the motivation for developing the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The package comprises two- and three-phase sampling regression estimators for global and small area estimations under simple and cluster sampling designs and cover 32 inventory scenarios in total. The estimators can be used for post-stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

The objective of this article is a) to establish the link between the mathematical formulation of the estimators and their implementation in our package, b) to illustrate their application by the respective functions in our package to real-world inventory scenarios and c) to highlight special cases, i.e. rare inventory scenarios, and demonstrate how the package-functions deal with such situations (including error-checking functions and data adjustments).

>>>>

% section:briefly give a review of existing R-packages for forest inventories (not much around) and 
Despite the broad range of methodological literature and case studies, freely available software supporting the application of the proposed multi-phase sampling methods in forest practice has been rare. Two exceptions are the \proglang{R} package \pkg{JoSAE} by \citet{josae2015} that provides the GREG \citep{sarndal2003} and EBLUP \citep{battese1988} two-phase small area estimator for simple sampling derived under the finite population approach, and the package \pkg{maSAE} by \citet{cullmann2016} who provided an initial implementation of the extended synthetic estimator under two and three-phase sampling \citep{mandallaz2013b, mandallaz2013c}. However, a more comprehensive software package covering a larger variety of sample designs and estimators for forest inventories has not yet been available. Our motivation has been to addressed this lack by the \proglang{R} package \pkg{forestinventory}. The package provides global and small area estimators for two-phase and three-phase forest inventories under simple and cluster sampling, which have been developed under the infinite population approach by Daniel Mandallaz at ETH Zurich between 2008 and 2017. The implemented methods have been demonstrated by case studies in Switzerland \citep{massey2014a, massey2015b, mandallaz2013b} and Germany \citep{hill2017a}. The implemented estimators cover 32 inventory scenarios and can be used for stratification, regression and regression within strata \citep{massey2015}. The long-term objective of \pkg{forestinventory} is to make the broad range of estimators available to a large user community and to facilitate their application in science as well as operational forest management.

The particular objective of this article is to \textbf{a)} establish the link between the mathematical description of the estimators and their implementation in our package, \textbf{b)} illustrate their application to real-world inventory data sets and \textbf{c)} adress special cases and demonstrate how the package-functions handle them.

### 2017-09-04 rnw/Main structure and modules of the Package.Rnw
update Andreas for alex_version 2
update Andreas 04.09.2017(5)

The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Figure \ref{fig:concmphase_and_sae_a}). The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$. The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e. the local density $Y(x)$ (e.g. the timber volume density per hectare). The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS). The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases (i.e. $\EX{(Y(x)-\hat{Y}(x))}$) using the subsample of terrestrial inventory units where the "true value" of $Y(x)$ is observed.  In the design-based context, the two-phase estimator is unbiased regardless of the model you use to produce the predictions.  This property comes from the assumption that each phase's sample is selected via simple random sampling. It should be noted that most forest inventories use systematic grids rather than simple random sampling, as is depicted in Figure \ref{fig:concmphase_and_sae_a}.  However, this is generally considered to be acceptable as long as the grid does not interact with periodic features in the forest structure (\cite{mandallaz2008} Chapter 7).
%In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.

@@

The selection of a suitable predictive model is an essential part in the application of multiphase inventories.  If the model is chosen poorly it could lead to inefficient use of the auxiliary information or even biased estimates.  This brings us to the two schools of thought concerning estimation: \textit{model-dependent} (also known as \textit{model-based}) and design-based. In the \textit{model-dependent} framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process. The model does not need to be fit from a probability sample, but it does need to be trusted in the sense that all of its model assumptions need to be verified.  For example, if we select a linear model using OLS then there is an implicit mathematical assumption that the residuals are \textit{iid}.  This model assumption needs to be empirically verified in order to avoid misinterpreting the regression coefficients and even the predictions. The unbiasedness of model-dependent estimators depend on the correct specification of the model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process. In contrast, the \textit{design-based} (also called \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$. At any given location $x_0$, the ground truth, $Y(x_0)$, is assumed to be fixed and \textit{not} the result of a stochastic process.  $Y(x)$ is random only because of the random selection of the plot locations $x$, which is determined by the sampling design.  A consequence of this approach is that the models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g. identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \cite{mandallaz2008}. The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics. They are therefore also referred to as \textit{model-assisted}. The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey. However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2008, mandallaz2016}.

@@

\begin{figure}[htb]
\centering
\resizebox{0.97\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}
\label{fig:struct_package}
\end{figure}

>>>>

The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly referred to as \textit{phases} (Figure \ref{fig:concmphase_and_sae_a}). The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$. The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e. the local density $Y(x)$ (e.g. the timber volume density per hectare). The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ using a linear regression model with ordinary least square (OLS). The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheaper predictions of $Y(x)$ in $x \in s_1$ and to correct any possible model biases, i.e. $\EX{(Y(x)-\hat{Y}(x))}$, using the subsample of terrestrial inventory units where the "true value" of $Y(x)$ is observed. In the design-based context, the two-phase estimator is unbiased regardless of the model you use to produce the predictions.  This property comes from the assumption that each phase's sample is selected via simple random sampling. It should be noted that most forest inventories use systematic grids rather than simple random sampling, as is depicted in Figure \ref{fig:concmphase_and_sae_a}.  However, this is generally considered to be acceptable as long as the grid does not interact with periodic features in the forest structure (\cite{mandallaz2008} Chapter 7).
%In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.

@@

The selection of a suitable predictive model is an essential part in the application of multiphase inventories.  If the model is chosen poorly it could lead to inefficient use of the auxiliary information or even biased estimates.  This brings us to the two schools of thought concerning estimation: \textit{model-dependent} (also known as \textit{model-based}) and design-based. In the \textit{model-dependent} framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process. The model does not need to be fit from a probability sample, but it does need to be trusted in the sense that all of its model assumptions need to be verified.  For example, if we select a linear model using OLS then there is an implicit mathematical assumption that the residuals are \textit{iid}.  This model assumption needs to be empirically verified in order to avoid misinterpreting the regression coefficients and even the predictions. The unbiasedness of model-dependent estimators depend on the correct specification of the model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process. In contrast, the \textit{design-based} approach rests upon the randomization of the sample locations $x$. At any given location $x$, the ground truth, $Y(x)$, is assumed to be fixed and \textit{not} the result of a stochastic process.  $Y(x)$ is random only because of the random selection of the plot locations $x$, which is determined by the sampling design.  A consequence of this approach is that the models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g. identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \cite{mandallaz2008}. The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics. They are therefore also referred to as \textit{model-assisted}. The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey. However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2008, mandallaz2016}.

@@

\begin{figure}[htb]
\centering
\resizebox{0.95\hsize}{!}{\includegraphics{fig/Package_structure_3.eps}}
\caption{Structure of the multi-phase estimators in the \proglang{R} package \pkg{forestinventory}}
\label{fig:struct_package}
\end{figure}
